---
title: "Homework 1: ggplot2"
author: "Piret Saar-Reismaa"
date: "2019-03-04"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(ggplot2)
```

By using *mpg* dataset:

1. Map a continuous variable to color, size, and shape. How do these aesthetics behave differently for categorical vs. continuous variables?f

Some aesthetics work with both, some not, giving different aesthetics to plotted points. 

- Color
```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, color = cty))
  
```

- Size
```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, size = cty))
```

- Shape
```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, shape = class))

```
Shape can not be mapped to a contiuos variable, it should have categorical values for each to have a certain shape. It will give an Error. 

2. What happens if you map the same variable to multiple aesthetics?

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, color = cty, size = cty))
```
Additional aesthetics can be used simultaniously to have one than more effect. 

3. What does the stroke aesthetic do? What shapes does it work with? (Hint: use ?geom_point)

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, shape = class), stroke=1, fill = "red", color = "green" )+
  scale_shape_manual(values = c(16:19, 21:24))
```
Stroke is used for shapes that have a border that can be colored separately, then stroe is used to modify the border width. 

4. What happens if you map an aesthetic to something other than a variable name, like aes(colour = displ < 5)?

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, color = displ<5))
```
Then it will differante between the given condition, so when displ is under 5 (displ<5) it is true and one color, and if not then a different color. 
